---
output: github_document
always_allow_html: true
---

# Hannah's Data Science Exam

The purpose of this ReadMe is to show my output and thinking from each question.


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
suppressMessages(library(tidyverse))
suppressMessages(list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.)))

```

# Question 1

In Question 1, COVID-19 cases and vaccinations are plotted across continents to display the differences between Africa and other continents. Secondly, to look at specific countries, the severity of the virus is examined in high and low life expectancy countries. General Hospitalisations and ICU admissions are plotted for continents and globally.

```{r}
suppressMessages(owid_covid_data <- read_csv("Q1/data/owid-covid-data.csv"))
source("Q1/code/plot_covid_and_vaccination.R")
plots <- plot_covid_and_vaccination(owid_covid_data)
plots$plot_cases  # Figure 1: Average COVID-19 Cases per Million by Continent over Time
plots$plot_vaccinations  # Figure 2: Average Fully Vaccinated per Hundred


suppressMessages(Deaths_by_cause <- read_csv("Q1/data/Deaths_by_cause.csv"))
combined_data <- bind_rows(owid_covid_data, Deaths_by_cause)

lowest_le <- combined_data %>%
filter(location %in% c("Chad", "Lesotho", "Central African Republic", "Nigeria", "Sierra Leone"))
Total_Deaths <- lowest_le$total_deaths

source("Q1/code/plot_lowest_le.R")
p <- plot_lowest_le(lowest_le, Total_Deaths, "Total Deaths over Time in Regions with Lowest LE")
p

highest_le <- combined_data %>%
filter(location %in% c("Hong Kong", "Japan", "Macao", "Monaco", "San Marino"))
Total_deaths <- highest_le$total_deaths
q <- plot_lowest_le(highest_le, Total_deaths, "Total Deaths over Time in Regions with Highest LE")
q

lowest_le <- combined_data %>%
filter(location %in% c("Chad", "Lesotho", "Central African Republic", "Nigeria", "Sierra Leone"))
Total_Cases_per_Million <- lowest_le$total_cases_per_million
r <- plot_lowest_le(lowest_le, Total_Cases_per_Million, "Total Cases per Million over Time in Regions with Lowest LE")
r


Total_cases_per_Million <- highest_le$total_cases_per_million

s <- plot_lowest_le(highest_le, Total_cases_per_Million, "Total Cases per Million over Time in Regions with Highest LE")
s

filtered_data <- combined_data %>%
  filter(continent %in% c("Africa", "North America", "South America", "Asia", "Europe"))
source("Q1/code/plot_filtered_data.R")
hosp <- plot_filtered_data(filtered_data, weekly_hosp_admissions_per_million, "Weekly Hospital Admissions Per Million", "Weekly Hospital Admissions among Continents")
hosp
icu <- plot_filtered_data(filtered_data, weekly_icu_admissions_per_million, "Weekly ICU Admissions Per Million", "Weekly ICU Admissions among Continents")
icu

source("Q1/code/plot_hospital_icu_data.R")
globalp <- plot_hospital_icu_data(combined_data, hosp_patients, icu_patients)
suppressWarnings(globalp)

```

# Question 2

In Question 2, London's weather is first examined, looking at monthly precipitation, cloudiness and coldness from 2019. Descriptive tables of weather trends are also provided.

The general UK weather is then examined, plotting average temperature and number of Heating Days per year since the 1880s. 

For London
```{r}

suppressMessages(london_weather <- read_csv("Q2/data/london_weather.csv"))
london_weather$date <- as.character(london_weather$date)
london_weather$date <- as.Date(london_weather$date, format = "%Y%m%d")
london <- na.omit(london_weather)
london$month <- format(london$date, "%Y-%m")
london_filtered <- london[london$date >= as.Date("2019-01-01"), ]

monthly_precipitation <- london_filtered %>%
  group_by(month) %>%
  summarize(total_precipitation = sum(precipitation, na.rm = TRUE))

monthly_cloudy <- london_filtered %>%
  group_by(month) %>%
  summarize(avg_cloud = mean(cloud_cover, na.rm = TRUE))

monthly_min_temp <- london_filtered %>%
  group_by(month) %>%
  summarize(min_temp = min(min_temp, na.rm = TRUE))

source("Q2/code/plot_monthly_data.R")
precip <- plot_monthly_data(monthly_precipitation, total_precipitation, " Total Precipitation")
precip

source("Q2/code/plot_monthly_data.R")
cloud <- plot_monthly_data(monthly_cloudy, avg_cloud, "Average Cloudiness")
cloud

source("Q2/code/plot_monthly_data.R")
cold <- plot_monthly_data(monthly_min_temp, min_temp, "Average Minimum Temp")
cold

```

```{r}
library(dplyr)
library(knitr)
library(kableExtra)

london$year <- lubridate::year(london$date)
source("Q2/code/generate_summary_stats.R")

summary_stats <- london %>% filter(london$date >= as.Date("2015-01-01"), ) %>%  group_by(year) %>%
  summarize(
    min_temp = min(min_temp, na.rm = TRUE),
    max_temp = max(max_temp, na.rm = TRUE),
    avg_temp = mean(mean_temp, na.rm =TRUE)


  )


kable(summary_stats, caption = "Stats Table", align = "c") %>%
    kable_styling()

source("Q2/code/generate_summary_stats_extra.R")
summary_stats_sun <- london %>% filter(london$date >= as.Date("2015-01-01"), ) %>%  group_by(year) %>%
  summarize(
    sunshine_mean = mean(sunshine, na.rm = TRUE),
    cloud_cover_mean = mean (cloud_cover, na.rm = TRUE),
    precipitation_mean = max(precipitation, na.rm = TRUE)
  )

kable(summary_stats_sun, caption = "Stats Table", align = "c") %>%
    kable_styling()


```


For the UK in general
```{r}

suppressMessages(UK <- read_csv("Q2/data/UKMonthly_Detailed.csv"))
UK$DATE <- as.Date(paste0(UK$DATE, "-01"), format = "%Y-%m-%d")

UK_yearly_temp <- UK %>%
  group_by(year = lubridate::year(DATE)) %>%
  summarize(average_temperature = mean(TAVG, na.rm = TRUE))

source("Q2/code/plot_yearly_data.R")
temp <- plot_yearly_data(UK_yearly_temp, average_temperature, "Average Temperature", "Yearly Average Temperature in UK")
temp


UK_yearly_heat <- UK %>%
  group_by(year = lubridate::year(DATE)) %>% summarize(average_heat = mean(HTDD, na.rm = TRUE))

heat <- plot_yearly_data(UK_yearly_temp, average_temperature, "Average Number of Heating Days", "Yearly Average Number of Heating Days in UK")
heat

```

# Question 3

In Question 3, Coldplay is examined first followed by Metallica. For both, average danceability per album and album popularity are plotted. I look at the correlation between danceability and popularity, finding similar results for both. When looking at energy levels for live vs studio albums, Coldplay gives substantially less energy in their studio albums.

I also look at the broader music scene and find the top 10 most danceable songs in recent history.

## Coldplay

```{r}
suppressMessages(Coldplay <- read_csv("Q3/data/Coldplay.csv"))

filtered_coldplay <- subset(Coldplay, album_name %in% c("Parachutes", "A Rush of Blood to the Head", "X&Y", "Viva La Vida or Death and All His Friends", "Mylo Xyloto", "Ghost Stories", "A Head Full of Dreams", "Love in Tokyo", "Everyday Life", "Music Of The Spheres"))

average_danceability <- aggregate(danceability ~ album_name, data = filtered_coldplay, FUN = mean)

source("Q3/code/plot_average_danceability.R")
suppressWarnings(dance_coldplay <- plot_average_danceability(average_danceability, album_name))
dance_coldplay


source("Q3/code/plot_popularity_band.R")
cold_play_plot <- plot_popularity_band(filtered_coldplay, album_name, "Popularity per ColdPlay Album")
cold_play_plot


source("Q3/code/create_corr.R")
col <- create_corr(filtered_coldplay, popularity, danceability, "Correlation between danceability and popularity")
col

#Looking at Coldplay's energy live vs studio
average_energy_cold <- Coldplay %>%
  filter(album_name %in% c("Ghost Stories", "Ghost Stories Live 2014")) %>% aggregate(energy ~ album_name, ., FUN = mean)

source("Q3/code/create_bar_plot.R")
bar_plot_cold <- create_bar_plot(average_energy_cold, album_name, energy, "Album", "Average Energy", "Average Energy for Coldplay Albums")
bar_plot_cold

```

## Metallica

```{r}
suppressMessages(metallica <- read.csv("Q3/data/metallica.csv"))
filtered_metallica <- subset(metallica, album %in% c("72 Seasons", "Metallica", "Master Of Puppets (Remastered)", "...And Justice For All", "Kill Em All (Remastered)", "Ride The Lightning (Remastered)", "Death Magnetic", "Garage Inc.", "HardwiredTo Self-Destruct", "Load"))
average_danceability_met <- aggregate(danceability ~ album, data = filtered_metallica, FUN = mean)

dance_met <- plot_average_danceability(average_danceability_met, album)
dance_met

met_plot <- plot_popularity_band(filtered_metallica, album, "Popularity per Metallica Album")
met_plot

met <- create_corr(filtered_metallica, popularity, danceability, "Correlation between danceability and popularity")
met

average_energy <- metallica %>%
filter(album %in% c("Some Kind Of Monster", "Some Kind Of Monster (Live)")) %>% aggregate(energy ~ album, ., FUN = mean)

bar_plot_met <- create_bar_plot(average_energy, album, energy, "Album", "Average Energy", "Average Energy for Metallica Albums")
bar_plot_met

```

## Broad Music Scene
```{r}
suppressMessages(Broader_Spotify_Info <- read.csv("Q3/data/Broader_Spotify_Info.csv"))

top_10_dance_songs <- Broader_Spotify_Info %>%
  top_n(10, danceability) %>%
  select(name, artist, danceability) %>%
  arrange(desc(danceability))

# top 10 dancing songs
table_dance <- kable(top_10_dance_songs, row.names = TRUE,
      caption = 'Top 10 Most Danceable Songs',
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T,
                      latex_options = c("striped",
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 13)
table_dance

```


# Question 4

Question 4 examines trends on Netflix. The top actors on Netflix according to their average IMDb rating are found, and their frequency across movies and series are tabled. This shows that more titles with these well-renowned actors need to shown on Netflix.

The most popular genres are also plotted according to TMBd popularity. Sci-Fi is the most popular, therefore Netflix should increase its Sci-Fi titles. The average run-time for the most popular movies and shows, respectively, is plotted to determine the optimal length of titles to maintain viewer interest. 

```{r}
library(knitr)
library(kableExtra)

suppressMessages(titles <- read_csv("Q4/data/titles.csv"))
suppressMessages(credits <- read_csv("Q4/data/credits.csv"))
merged_data <- merge(titles, credits, by = "id", all.x = TRUE)

source("Q4/code/calculate_actor_frequency.R")
actor_frequency <- calculate_actor_frequency(merged_data)

table <- kable(actor_frequency, row.names = TRUE,
      caption = 'Top Actors on Netflix by IMDB Score and their Frequency in Netflix titles',
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T,
                      latex_options = c("striped",
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 13)
table

source("Q4/code/genre_plot.R")
genre <- genre_plot(merged_data)
genre

source("Q4/code/plot_runtime_distribution.R")
dist_movie <- plot_runtime_distribution(merged_data, "MOVIE")
#dist_movie

dist_series <- plot_runtime_distribution(merged_data, "SHOW")
#dist_series

```

# Question 5

To research ideas for a new app, apps are grouped by categories and arranged according to their average ratings. The top ten categories are then presented.

The number of apps per category is further plotted as well as their number of installations, showing which categories are over-saturated and which receive the highest number of installations.

```{r}

suppressMessages(googleplaystore <- read_csv("Q5/data/googleplaystore.csv"))
googleplaystore$Installs <- factor(googleplaystore$Installs, levels = c("1,000,000,000+", "500,000,000+", "100,000,000+", "50,000,000+", "10,000,000+", "5,000,000+", "1,000,000+", "500,000+", "100,000+", "50,000+", "10,000+", "5,000+", "1,000+", "500+", "100+", "50+", "10+", "5+", "1+", "0+", "0"), ordered = TRUE)


category_avg_ratings <- googleplaystore %>% group_by(Category) %>%
  summarize(Average_Rating = mean(Rating, na.rm = TRUE)) %>% top_n(10, Average_Rating) %>% arrange(desc(Average_Rating))

table_app <- kable(category_avg_ratings, row.names = TRUE,
      caption = 'Top 10 App Categories by Average Rating',
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T,
                      latex_options = c("striped",
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 13)
table_app


```


```{r, warning = FALSE, fig.align = 'center', fig.ext = "png", fig.height = 8, fig.width = 7}
source("Q5/code/plot_app_installs.R")
installp <- plot_app_installs(googleplaystore)
installp
```

